---
title: "R basics"
author: Katie Schuler
subtitle: "Data Science for Studying Language and the Mind"
date: 2024-08-29
echo: true
format: 
    revealjs:
        theme: dark
        incremental: true
        footer: "[https://kathrynschuler.com/datasci](/index.html)"

---

## Announcements

- Course is full
- Thursday 3:30 moved to 5:15pm 
- Please join the waitlist 
- Please fill out Who's in class 

## You are `here` {.smaller} 

:::: {.columns}

::: {.column width="33%"}

##### Data science with R 
::: {.nonincremental}
- `R basics`
- Data visualization 
- Data wrangling 
:::
:::

::: {.column width="33%"}

##### Stats & Model buidling
::: {.nonincremental}
- Sampling distribution
- Hypothesis testing
- Model specification
- Model fitting 
- Model accuracy
- Model reliability
:::
:::

::: {.column width="33%"}

##### More advanced 
::: {.nonincremental}

- Classification
- Inference for regression
- Mixed-effect models
::: 
:::

::::

# Today

Let's find what we need for today (see course website)

## Basic concepts {.smaller}

- `Expressions`: fundamental building blocks of programming 
- `Objects`: allow us to store stuff, created with assignment operator
- `Names`: names w give objects must be letters, numbers, ., or _
- `Attributes`: allow us to attach arbitrary metadata to objects
- `Functions`: take some input, perform some computation, and return some output
- `Environment`: collection of all objects we defined in current R session
- `Packages`: collections of functions, data, and documentation bundled together in R
- `Comments`: notes you leave for yourself, not evaluated
- `Messages`: notes R leaves for you (FYI, warning, error)


## Important functions 
#### Objects
- `str(x)` - returns summary of object's structure
- `typeof(x)` - returns object's data type
- `length(x)` - returns object's length
- `attributes(x)` - returns list of object's attributes

## Important functions  
#### Environment

- `ls()` - list all variables in environment
- `rm(x)` - remove x variable from environment
- `rm(list = ls())` - remove all variables from environment

## Important function 
#### Packages

- `install.packages()` to install packages
- `library()` to load package into current R session.
- `data()` to load data from package into environment
- `sessionInfo()` - version info, packages for current R session 

## Important functions  
####  Help

- `?mean` - get help with a function
- `help('mean')` - search help files for word or phrase
- `help(package='tidyverse')` - find help for a package 

# Vectors


## Vectors

are fundamental data structures in R. There are two types:

- **atomic vectors** - elements of the same data type
- **lists** - elements refer to any object 




## Atomic vectors 

Atomic vectors can be one of six **data types**:

|`typeof(x)` | examples |
|-|---|
| double | 3, 3.32 |
| integer | 1L, 144L |
| character | "hello", 'hello, world!'|
| logical | TRUE, F |

. . . 

*atomic* because they must contain only one type

## Atomic vectors 

:::: {.columns}

::: {.column width="50%"}

`double` 

```{r}
typeof(3.34)
```

`integer` 

```{r}
typeof(3L)
```


:::

::: {.column width="50%"}
`character` 

```{r}
typeof('hello, world!')
```

`logical` 

```{r}
typeof(TRUE)
```

:::

::::



## Create a vector


with `c()` for `concatenate`

```{r}

c(2,4,6)
c("hello", "world", "!")
c(T, F, T)
c("hello", c(2, 3))
```


## Create a vector

with sequences `seq()` or repetitions `rep()`

```{r}
# sequence of integers have a special shorthand
6:10
```

. . . 

```{r}
# sequence from, to, by 
seq(from=3, to=5, by=0.5)
```

. . . 

```{r}
# rep(x, times = 1, each = 1)
rep(c(1,0), times = 4)
```

. . . 

```{r}
# rep(x, times = 1, each = 1)
rep(c(1,0), each = 4)
```

## Check data type

with `typeof(x)` - returns the type of vector x

```{r}
typeof(3)
typeof(3L)
typeof("three")
typeof(TRUE)

```

## Check data type

with `is.*(x)` - returns `TRUE` if x has type `*`

```{r}
is.double(3)
is.integer(3L)
is.character("three")
is.logical(TRUE)
```

## Coercion, implicit

If you try to include elements of different types, R will coerce them into the same type without warning (**implicit coercion**)

```{r}
x <- c(1, 2, "three", 4, 5 )
x
typeof(x)
```

## Coercion, explicit

You can also use **explict coercion** to change a vector to another data type with `as.*()`

```{r}
x <- c(1, 0 , 1, 0)
as.logical(x)
```

. . . 

::: aside
hierarchy: character > double > integer > logical 
:::

# More complex structures 

## More complex structures {.smaller}

Some more complex data structures are **built from atomic vectors** by adding **attributes**:

| Structure | Description | 
| -- | -------- |
| `matrix` | vector with `dim` attribute representing 2 dimensions
| `array` | vector with `dim` attribute representing n dimensions
| `data.frame` | a named list of vectors (of equal length) with attributes for `names` (column names), `row.names`, and `class="data.frame"` |

## Create more complex structures 


:::: {.columns}

::: {.column width="50%"}

`matrix` 

```{r}
matrix(0, nrow=2, ncol=3)
```




`data.frame` 

```{r}
data.frame(x=c(1,2,3), y=c('a','b','c'))
```


:::

::: {.column width="50%"}

`array` 

```{r}
array(0, dim=c(2,3,2))
```


:::

::::

# Operations

## Basic math operators


| Operator | Operation |
| -- | -- |
| `()` | Parentheses | 
| `^` | Exponent | 
| `*` | Multiply | 
| `/` | Divide | 
| `+` | Add |
 `-` | Subtract |


## Basic math operations

follow the order of operations you expect (PEMDAS)


```{r}
# multiplication takes precedence
2 + 3 * 10

# we can use paratheses to be explicit
(2 + 3) * 10 
```




## Comparison operators

| Operator | Comparison | 
| -- | -- |
| `x < y` | less than |
| `x > y` | greater than | 
| `x <= y` | less than or equal to | 
| `x >= y` | greater than or equal to |
| `x != y` | not equal to | 
| `x == y` | equal to | 

## Comparison operators


```{r}
x <- 2
y <- 3
```

. . . 

<br> 

```{r}
x < y
x > y 
x != y
x == y
```

## Logical operators 

| Operator | Operation | 
| -- | -- | 
| `x | y` | or |
| `x & y` | and |
| `!x` | not |
| `any()` | true if any element meets condition | 
|`all()` | true if all elements meet condition | 
| `%in%` | true if any element is in following vector | 

## Logical operators 

```{r}
x <- TRUE
y <- FALSE
```

. . . 

<br> 

```{r}
x | y
x & y 
!x 
any(c(x,y))
all(c(x,y))
```

## Operations are vectorized 

Almost all operations (and many functions) are vectorized 

. . .

:::: {.columns}

::: {.column width="50%"}

`math` 

```{r}
c(1, 2, 3) + c(4, 5, 6)
c(1, 2, 3) / c(4, 5, 6)
c(1, 2, 3) * 10 
c(1, 2, 30) > 10
```



:::

::: {.column width="50%"}
`logical` 

```{r}
x <- c(TRUE, FALSE, FALSE)
y <- c(TRUE, TRUE, FALSE)
z <- TRUE
```

```{r}
x | y
x & y 
x | z 
x & z 
```

:::

::::



## Operator coercion 

Operators and functions will also coerce values when needed (and without warning)

```{r}
5.6 + 2L
10 + FALSE 
log(1)
log(TRUE)
```




# Subsetting 

## Subsetting

> Subsetting is a natural complement to str(). While str() shows you all the pieces of any object (its structure), subsetting allows you to pull out the pieces that you’re interested in. ~ Hadley Wickham, Advanced R

`str()`

```{r}
x <- c("hello", "world", "!")
str(x)

y <- c(1, 2, 3, 4, 5)
str(y)
```

## Subsetting

There are three operators for subsetting objects: 

- `[` - *subsets* (one or more) elements
- `[[` and `$` - *extracts* a single element 

---
## Subset multiple elements with `[` {.smaller}

| Code | Returns |
| ---- | -------- |
| `x[c(1,2)]` | positive integers select elements at specified indexes |
| `x[-c(1,2)]` | negative integers select all but elements at specified indexes |
|`x[c("x", "y")]` | select elements by name, if elements are named |
| `x[]` |  nothing returns the original object |
| `x[0]` | zero returns a zero-length vector |
| `x[c(TRUE, TRUE)]` | select elements where corresponding logical value is TRUE |

## Subset multiple elements with `[` {.smaller}


:::: {.columns}

::: {.column width="50%"}

`atomic vector`

```{r}
x <- c("hello", "world", "1")
```

```{r}
x[c(1,2)]
x[-c(1,2)]
x[]

```




:::

::: {.column width="50%"}

`data.frame`

```{r}
y <- data.frame(
        this = c(1, 2,3), 
        that = c("a", "b", "c"),
        theother = c(4, 5, 6)
        )
```

```{r}

y[c(1,2)]
y[-c(1,2)]
y[c("this")]

```


:::

:::: 



## 3 ways to extract a single element {.smaller}

| Code | Returns |
| ---- | -------- |
| `[[2]]` | a single positive integer (index) |
| `[['name']]` | a single string  |
| `x$name` | the `$` operator is a useful shorthand for `[['name']]` |

## 3 ways to extract a single element {.smaller}


:::: {.columns}

::: {.column width="50%"}

`atomic vector`

```{r}
x <- c("hello", "world", "1")
```

```{r}
x[[1]]
x[[2]]
x[[3]]

```




:::

::: {.column width="50%"}

`data.frame`

```{r}
y <- data.frame(
        this = c(1, 2,3), 
        that = c("a", "b", "c"),
        theother = c(4, 5, 6)
        )
```

```{r}

y[[1]]
y[["that"]]
y$that

```


:::

:::: 



## R has many built-in functions

```{r}
x <- c(1, -2, 3)
```

. . . 

:::: {.columns}

::: {.column width="50%"}

Some are vectorized

```{r}
log(x)
abs(x)
round(x, 2)

```


:::

::: {.column width="50%"}

Some are not

```{r}
mean(x)
max(x)
min(x)
```


:::

:::: 






## Missing values {.smaller}

:::: {.columns}

::: {.column width="50%"}

`NA`

- used to represent missing or unknown elements in vectors
- Note that `NA` is contageous: expressions including `NA` usually return `NA`
- Check for `NA` values with `is.na()`

```{r}
x <- c(1, NA, 3)
is.na(x)
length(x)
mean(x)
```



:::

::: {.column width="50%"}

`NULL` 

- used to represent an empty or absent vector of arbitrary type
- `NULL` is its own special type and always has length zero and `NULL` attributes
- Check for `NULL` values with `is.null()`

```{r}
x <- c()
is.null(x)
length(x)
mean(x)
```


:::

::::





## Programming {.smaller}


:::: {.columns}

::: {.column width="50%"}

`functions`

are reusable pieces of code that take some input, perform some task or computation, and return an output


```r
function(inputs){
    # do something
    return(output)
}
```


:::

::: {.column width="50%"}

`control flow` 

refers to managing the order in which expressions are executed in a program

- `if`...`else` - if something is true, do this; otherwise do that
- `for` loops - repeat code a specific number of times 
- `while` loops - repeat code as long as certain conditions are true
- `break` - exit a loop early
- `next` - skip to next iteration in a loop 

:::

::::

# Subsetting quirks 

If we have time 

## Notes on `[` with higher dim objects 

```{r}
m <- matrix(1:6, nrow=2, ncol=3)
m
```

```{r}
# separate dimensions by comma 
m[1, 2]

# omitted dim return all from that dim 
m[2, ]
m[ , 2]

```

## Notes on `[[` and `$`: {.smaller}

both `[[` and `[` work for vectors; use `[[`

```{r}
x <- c(1, -2, 3)
x[[1]]
x[1]
```

`$` does partial matching without warning

```{r}
df <- data.frame(
        this = c(1, 2,3), 
        that = c("a", "b", "c"),
        theother = c(4, 5, 6)
        )
```

```{r}
df[['theo']]
df$theo
```

# Questions?

Have a great weekend!