---
title:  "Lab 5: Hypothesis testing"
subtitle: "Not graded, just practice"
date:   2024-09-26
author: Katie Schuler
---

Practice your new stats skills with these practice exam questions! Best to open a fresh Google Colab notebook and test things out! Refer to the study guide to find answers as well. 

::: {.callout-tip title="Tip"}
More than one answer may be correct!  
:::


```{r}
#| echo: false
#| message: false

library(webexercises)
library(tidyverse)
```

If you would like to practice with a set of data, you can import the following dataset with `read_csv`. Note that females are coded as NA in this dataset! 

```r
# brain volumes simulated from Ritchie et al
"http://kathrynschuler.com/datasets/brain_volume.csv"
```


## Visualize a categorical variable

(a) Which of the following is the best choice to visualize a categorical variable? Choose one.


    ```{r}
    #| echo: false
    #| results: asis

    # Define the answer choices
    choices <- c(
    "`geom_rug()`", "`geom_histogram()`", answer="`geom_boxplot()`", "`geom_point()`", "`geom_smooth()`"
    )

    cat(longmcq(choices))
    ```

(b) Which of the following figures shows a box plot? 

    ```{r}
    #| echo: false
    #| layout-ncol: 2
    theme_set(theme_minimal(base_size = 15))
    data <- read_csv('http://kathrynschuler.com/datasets/brain_volume.csv') %>%
        mutate(sex = ifelse(sex == "male", sex, "female")) 

    ggplot(data, aes(x = sex, y = volume, color =sex)) +
        geom_boxplot() 
    ggplot(data, aes(x = sex, y = volume, color = sex)) +
        geom_bar(stat = "identity") 
    ```

    ```{r}
    #| echo: false
    #| results: asis

    # Define the answer choices
    choices <- c(
    answer = "left", 
    "right", 
    "both the left and right", 
    "neither the left and right"
    )

    cat(longmcq(choices))
    ```

(c) True or false, the difference in means between male and female participants in our sample is subject to sampling variability? 

    ```{r}
    #| echo: false
    #| results: asis

    # Define the answer choices
    choices <- c(
    answer ="True", 
    "False"
    )

    cat(longmcq(choices))
    ```

## Hypothesis testing framework 

(a) Fill in the blanks about the 3-step hypothesis testing framework: (1) First we pose a  `r fitb("null hypothesis", width = 10)`, (2) then we ask if true, how likly is our observed pattern of results? This likelihood is quanitfied with a (`r fitb("p-value", width = 10)`), and (3) finally if the liklihood is leass than some thershold, we (`r fitb("reject", width = 10)`) the null hypothesis. 


(b) What is the practical reason we pose a null hypothesis? 

    ```{r}
    #| echo: false
    #| results: asis

    # Define the answer choices
    choices <- c("It is the hypothesis most likely to be true.",
                "It allows us to generate predictions based on prior beliefs.",
                answer = "It is the hypothesis for which we can simulate data.",
                "It ensures that the alternative hypothesis is proven false.")


    cat(longmcq(choices))
    ``` 

(c) True or false, Randomization simulates a world in which there is no relationship between brain volume and sex. 

    ```{r}
    #| echo: false
    #| results: asis

    # Define the answer choices
    choices <- c(
        answer = "True", 
        "False"
    )


    cat(longmcq(choices))
    ``` 

(d) Which of the following would compute a p-value? 

    ```{r}
    #| echo: false
    #| results: asis

    # Define the answer choices
    choices <- c(
        answer = "Count the number of values in our null distribution that are more extreme than our observed value.", 
        answer = "Call `get_p_value()` on an infer object in which we have constructed a null distribution.",
        "p-values are not computed.", 
        "Add up all of the differences in means"
        ) 
        


    cat(longmcq(choices))
    ``` 

(e) A large p-value means our observed value is very  `r fitb("likely", width = 10)` under the null hypothesis. A small p-value means our observed value is very `r fitb("unlikely", width = 10)` under the null hypothesis. 

(f) True or false, a p-value of less than 0.05 indicates obtaining our observed value under the null is *impossible*

    ```{r}
    #| echo: false
    #| results: asis

    # Define the answer choices
    choices <- c(
        "True", 
        answer = "False"
    )


    cat(longmcq(choices))
    ``` 

(g) A type I error is also known as a `r fitb("false positive", width = 10)`(wrongly thinking that the effect is present); a type II error is also known as a `r fitb("false negative", width = 10)` (wrongly thikning the effect is absent)

## There is only one test 

(a) There are two ways we can construct a sampling distribution: (1)  `r fitb("nonparametrically", width = 10)`, via brute computational force; and  (2) `r fitb("parametrically", width = 10)`, by assuming the data were sampled from a known probability distribution. 

(b) True or false, we can use the `t.test()` function or `assume(distribution = "t")` with an infer object to calcuate a p-value for a t-test. 

    ```{r}
    #| echo: false
    #| results: asis

    # Define the answer choices
    choices <- c(
       answer = "True", 
        "False"
    )


    cat(longmcq(choices))
    ``` 

## Correlation 

(a) We can explore the relationship between two quantities visually with a scatter plot. Which geom is best suited for this? 
    ```{r}
    #| echo: false
    #| results: asis

    # Define the answer choices
    choices <- c(
        "`geom_histogram()`", 
        "`geom_boxplot()`", 
        answer ="`geom_point()`", 
        "`geom_dotplot()`",
        "`geom_density()`"
    )


    cat(longmcq(choices))
    ``` 

(b) If there is no relationship between variables, we say they are `r fitb("independent", width = 10)`

(c) One way to quantify `r fitb("linear", width = 10)` relationships is with correlation. 

(d) Correlation ranges from -1 to 1, where: 

    ```{r}
    #| echo: false
    #| results: asis

    # Define the answer choices
    choices <- c(
        "-1 means mostly negative and 1 means mostly positive",
        answer = "-1 means perfectly negative and 1 means pefectly positive",
        "-1 means low correlation and 1 means high correlation"
    )


    cat(longmcq(choices))
    ``` 

(e) True or false, on an infer object, we can calcuate the correlation with `calcuate(stat = "correlation")`

    ```{r}
    #| echo: false
    #| results: asis

    # Define the answer choices
    choices <- c(
    answer = "True", 
        "False"
    )


    cat(longmcq(choices))
    ``` 


(f) Suppose we quantify how likely our observed correlation is under the null hypothesis and our p-value is 0.68. Should we reject the null hypothesis? 


    ```{r}
    #| echo: false
    #| results: asis

    # Define the answer choices
    choices <- c(
    "Yes", 
    answer = "No"
    )


    cat(longmcq(choices))
    ``` 

