---
title: "Applied model specification"
author: 
    - Katie Schuler
date: 2024-10-08
---

::: {.callout-warning title="Under construction"}

Still working on this!
:::

Now that we've covered the terminology and concepts, let's apply model specification to some real models. 

```{r}
#| message: false

library(tidyverse)
library(mosaic)
```


## "Toy" data 


Let's start with the simplest possible example, a dataset with two data points. Suppose you record how many days you study over the next two days. On day 1, you study for 2 hours. On day 2, you study for 3 hours. Your dataset might look something like this. 

::: {.panel-tabset}

## Plot



```{r}
#| echo: false
#| fig-height: 3
#| fig-width: 3

toy_data <- tibble(
  x = c(1, 2, 3, 4, 5), 
  y = c(2, 6, 7, 12, 13)
)

toy_data %>%
  ggplot(aes(
    x = x, 
    y = y
  )) +
  geom_point() +
  theme_bw(base_size = 14) 

```


## Data

```{r}
#| echo: false
#| layout-ncol: 2


library(knitr)
model
```


## Code

```r
toy_data <- tibble(
  x = c(1, 2, 3, 4, 5), 
  y = c(2, 6, 7, 12, 13)
)

toy_data %>%
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  theme_bw(base_size = 14) 

```
:::

1. **Specify our response variable**, $y$: the response variable (🥸 data, output, prediction) is the variable you are trying to predict or explain with your model.
    - `y`

2. **Specify explantory variables**, $x_i$: the explanatory variables (🥸 regressors, inputs, predictors) are the predictors in your data that could help explain the response variable. Our data has only one possible: 
    - `x` 

3. **Specify the functional form**: the functional form describes the relationship between the response and explanatory variables with a mathematical expresson. In a linear model, we express this relationship as a weighted sum of inputs: 
    - $y=\sum_{i=1}^{n}w_ix_i$

4. **Specify model terms**: here we need to specify exactly *how* to express our explanatory variables in our functional form. The actual variables and constants that will be included in the model. There are four kinds of terms: (1) intercept, (2) main, (3) interaction, and (4) transformation. Here we have the simplest case of an intercept and one main term (no interactions or transformations necessary)
    - $y = w_1\mathbf{1} + w_2x_2$
    - in R: `y ~ 1 + x`

::: {.panel-tabset}

## Plot

:::: {.columns}
::: {.column width="50%"}


```{r}
#| echo: false
#| fig-height: 3
#| fig-width: 3

model <- lm(y ~ 1 + x, data = toy_data)

toy_data <- toy_data %>%
  mutate(w_formula = -0.4*1 + 2.8*x) %>%
  mutate(w_predict= predict(model, toy_data))

toy_data %>%
  ggplot(aes(
    x = x, 
    y = y
  )) +
  geom_point() +
  geom_line(aes(y = w_predict), color = "blue") +
  theme_bw(base_size = 14) 

```

:::
::: {.column width="50%"}

Model specification: <br>
$y = w_1\cdot\mathbf{1} + w_2\cdot\mathbf{x}$

```{r}
#| echo: false

lm(y ~ 1 + x, data = toy_data)
```

Fitted model: <br>
$y = -0.4\cdot1 + 2.8\cdot x$

:::
::::


## Data

```{r}
#| echo: false
#| layout-ncol: 2


model
```


## Code

```r
model <- lm(y ~ 1 + x, data = toy_data)

toy_data <- toy_data %>%
  mutate(with_formula = -0.4*1 + 2.8*x) %>%
  mutate(with_predict= predict(model, toy_data))

toy_data %>%
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_line(aes(y = with_predict), color = "blue") +
  theme_bw(base_size = 14) 

```
:::

## Swim records

### One input


::: {.panel-tabset}

## Plot

:::: {.columns}
::: {.column width="50%"}


```{r}
#| echo: false
#| fig-height: 3
#| fig-width: 3

model <- lm(time ~ 1, data = SwimRecords)

SwimRecords_predict <- SwimRecords %>%
  mutate(with_formula = 59.92*1) %>%
  mutate(with_predict= predict(model, SwimRecords))

SwimRecords_predict %>%
  ggplot(aes(x = year, y = time)) +
  geom_point() +
  geom_line(aes(y = with_predict), color = "blue") +
  theme_bw(base_size = 14) 

```

:::
::: {.column width="50%"}

Model specification: <br>
$y = w_1\cdot\mathbf{1}$

```{r}
#| echo: false

model
```

Fitted model: <br>
$y = (59.92)1$

:::
::::


## Data

```{r}
#| echo: false
#| layout-ncol: 2


library(knitr)
kable(SwimRecords_predict)
```


## Code

```r
model <- lm(time ~ 1, data = SwimRecords)

SwimRecords_predict <- SwimRecords %>%
  mutate(with_formula = 59.92*1) %>% 
  mutate(with_predict= predict(model, SwimRecords))

SwimRecords_predict %>%
  ggplot(aes(x = year, y = time)) +
  geom_point() +
  geom_line(aes(y = with_predict), color = "blue") +
  theme_bw(base_size = 14) 

```
:::


### Two inputs


::: {.panel-tabset}

## Plot

:::: {.columns}
::: {.column width="50%"}


```{r}
#| echo: false
#| fig-height: 3
#| fig-width: 3

model <- lm(time ~ 1 + year, data = SwimRecords)

SwimRecords_predict <- SwimRecords %>%
  mutate(with_formula = 567.2420*1 + -0.2599*year) %>%
  mutate(with_predict= predict(model, SwimRecords))

SwimRecords_predict %>%
  ggplot(aes(x = year, y = time)) +
  geom_point() +
  geom_line(aes(y = with_predict), color = "blue") +
  theme_bw(base_size = 14) 

```

:::
::: {.column width="50%"}

Model specification: <br>
$y = w_1\cdot\mathbf{1} + w_2\cdot\mathbf{year}$

```{r}
#| echo: false

model
```

Fitted model: <br>
$y = (567.2420)\mathbf{1} + (-0.2599)\mathbf{year}$

:::
::::


## Data

```{r}
#| echo: false
#| layout-ncol: 2


library(knitr)
kable(SwimRecords_predict)
```


## Code

```r
model <- lm(time ~ 1 + year, data = SwimRecords)

SwimRecords_predict <- SwimRecords %>%
  mutate(with_formula = 567.2420*1 + -0.2599*year) %>%
  mutate(with_predict= predict(model, SwimRecords))

SwimRecords_predict %>%
  ggplot(aes(x = year, y = time)) +
  geom_point() +
  geom_line(aes(y = with_predict), color = "blue") +
  theme_bw(base_size = 14) 

```
:::


### Three inputs


::: {.panel-tabset}

## Plot

:::: {.columns}
::: {.column width="40%"}


```{r}
#| echo: false
#| fig-height: 3
#| fig-width: 3

model <- lm(time ~ 1 + year + sex, data = SwimRecords)

SwimRecords_predict <- SwimRecords %>%
  mutate(with_formula = 555.7168*1 + -0.2515*year + -9.7980 *sex) %>%
  mutate(with_predict= predict(model, SwimRecords))

SwimRecords_predict %>%
  ggplot(aes(x = year, y = time, shape = sex)) +
  geom_point() +
  geom_line(aes(y = with_predict), color = "blue") +
  theme_bw(base_size = 14) 

```

:::
::: {.column width="60%"}

Model specification: <br>
$y = w_1\cdot\mathbf{1} + w_2\cdot\mathbf{year} + w_3\cdot\mathbf{sex}$

```{r}
#| echo: false

model
```

Fitted model: <br>
$y = (555.7168)\mathbf{1} + (-0.2515 )\mathbf{year} + (-9.7980)\mathbf{sex}$

:::
::::


## Data

```{r}
#| echo: false
#| layout-ncol: 2


library(knitr)
kable(SwimRecords_predict)
```


## Code

```r
model <- lm(time ~ 1 + year, data = SwimRecords)

SwimRecords_predict <- SwimRecords %>%
  mutate(with_formula = 555.7168*1 + -0.2515*year + -9.7980 *sex) %>%
  mutate(with_predict= predict(model, SwimRecords))

SwimRecords_predict %>%
  ggplot(aes(x = year, y = time)) +
  geom_point() +
  geom_line(aes(y = with_predict), color = "blue") +
  theme_bw(base_size = 14) 

```
:::


### Interaction 

::: {.panel-tabset}

## Plot

:::: {.columns}
::: {.column width="40%"}


```{r}
#| echo: false
#| fig-height: 3
#| fig-width: 3

model <- lm(time ~ 1 + year * sex, data = SwimRecords)

levels(SwimRecords$sex)
levels(as.numeric(SwimRecords$sex)[0])


SwimRecords_predict <- SwimRecords %>%
  mutate(with_formula = 697.3012*1 + -0.3240*year + -302.4638*as.numeric(sex) + 0.1499*year*as.numeric(sex)
    ) %>%
  mutate(with_predict= predict(model, SwimRecords))

SwimRecords_predict %>%
  ggplot(aes(x = year, y = time, shape = sex)) +
  geom_point() +
  geom_line(aes(y = with_predict), color = "blue") +
  theme_bw(base_size = 14) 

```

:::
::: {.column width="60%"}

Model specification: <br>
$y = w_1\cdot\mathbf{1} + w_2\cdot\mathbf{year} + w_3\cdot\mathbf{sex} + w_4\cdot\mathbf{year\times{sex}}$

```{r}
#| echo: false

model
```

Fitted model: <br>
$y = (697.3012)\mathbf{1} + (-0.3240)\mathbf{year} + (-302.4638)\mathbf{sex} + (0.1499)\mathbf{year\times{sex}}$

:::
::::


## Data

```{r}
#| echo: false
#| layout-ncol: 2


library(knitr)
kable(SwimRecords_predict)
```


## Code

```r
model <- lm(time ~ 1 + year, data = SwimRecords)
as.numeric(SwimRecords$sex)
levels(SwimRecords$sex)

SwimRecords_predict <- SwimRecords %>%
  mutate(with_formula = 697.3012*1 + -0.3240*year +
    -302.4638*as.numeric(sex) + 0.1499*(year*as.numeric(sex))
    ) %>%
  mutate(with_predict= predict(model, SwimRecords))

SwimRecords_predict %>%
  ggplot(aes(x = year, y = time)) +
  geom_point() +
  geom_line(aes(y = with_predict), color = "blue") +
  theme_bw(base_size = 14) 

```
:::


### Transformation 

::: {.panel-tabset}

## Plot

:::: {.columns}
::: {.column width="40%"}


```{r}
#| echo: false
#| fig-height: 3
#| fig-width: 3

model <- lm(time ~ 1 + year * sex +I(year^2), data = SwimRecords)

SwimRecords_predict <- SwimRecords %>%
  mutate(with_formula = 567.2420*1 + -0.2599*year) %>%
  mutate(with_predict= predict(model, SwimRecords))

SwimRecords_predict %>%
  ggplot(aes(x = year, y = time, shape = sex)) +
  geom_point() +
  geom_line(aes(y = with_predict), color = "blue") +
  theme_bw(base_size = 14) 

```

:::
::: {.column width="60%"}

Model specification: <br>
$$
y = w_1\cdot\mathbf{1} + w_2\cdot\mathbf{year} + w_3\cdot\mathbf{sex}+ w_4\cdot\mathbf{year\times{sex}} +w_5\cdot\mathbf{year}^2
$$

```{r}
#| echo: false

model
```

Fitted model: <br>
\begin{align}
y = &(1.110e+04)\mathbf{1} + (-1.098e+01)\mathbf{year} + (-3.171e+02)\mathbf{sex}\\
&+ (1.575e-01 )\mathbf{year\times{sex}} + (2.729e-03)\mathbf{year}^2
\end{align}

:::
::::


## Data

```{r}
#| echo: false
#| layout-ncol: 2


library(knitr)
kable(SwimRecords_predict)
```


## Code

```r
model <- lm(time ~ 1 + year, data = SwimRecords)

SwimRecords_predict <- SwimRecords %>%
  mutate(with_formula = 567.2420*1 + -0.2599*year) %>%
  mutate(with_predict= predict(model, SwimRecords))

SwimRecords_predict %>%
  ggplot(aes(x = year, y = time)) +
  geom_point() +
  geom_line(aes(y = with_predict), color = "blue") +
  theme_bw(base_size = 14) 

```
:::



## Brain size (log)

## Plant heights (polynomials)







## Further reading

- [Ch 6: Language of models](https://dtkaplan.github.io/SM2-bookdown/language-of-models.html) in Statistical Modeling

